### 2022-02-24 documentation/paper.tex
Starting documentation

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\begin{document}
\begin{titlepage}
\centering
{\scshape\LARGE University of Potsdam \par}
\vspace{1cm}
{\scshape\Large Final Thesis for a Bachelor's Degree in Computational Science\par}
\vspace{1.5cm}
{\huge\bfseries Map Abstraction for Multi-Agent Pathfinding problems with Answer Set Programming\par}
\vspace{2cm}
{\Large\itshape Adrian Salewsky\par}
\vfill
supervised by\par 
\vspace{\baselineskip}
Prof Torsten Schaub \\
\vspace{\baselineskip}
Arvid Becker
\vfill
{01.04.2022 \par}

\begin{abstract}
% Abstrakt auf deutsch
% MAPF-Probleme
% Ziel: Abstraktion
% warum abstrahieren
% warum ASP
\end{abstract}
\end{titlepage}

\section{Introduction}
% abstract 
% ???

\section{Asprilo}
% what is asprilo
% M-domain
% changes in goal condition to compare

\section{Abstraction Methods}
\subsection{Shortest Path}
\subsection{Node Combining}
\subsection{Reachable Nodes}

\section{Benchmarking}

\section{Conclusion}

\newpage

\begin{thebibliography} {0}

\end{thebibliography}

\newpage
\appendix
\newpage
\section{Affidavit}
I hereby affirm that this Bachelor's Thesis represents my own written work and that I have used no source and aids other than those indicated.
All passages quoted from publications or paraphrased from these sources are properly cited and attributed.
\vspace{\baselineskip}
\vspace{\baselineskip}
\vspace{\baselineskip}
\vspace{\baselineskip}

\noindent\begin{tabular}{lll}
\makebox[5cm]{\hrulefill} & \hspace{2cm} & \makebox[5cm]{\hrulefill}\\
Date, Place &  & Signature

\end{tabular}

\end{document}