### 2022-03-31 presentation in progress/slides.tex
Creating Presentation

\documentclass{beamer}

\usepackage[utf8]{inputenc}
\usepackage{listings}
\usetheme{Copenhagen}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{headline}{}
\setbeamertemplate{footline}{}

\title{Map Abstraction for Multi-Agent Pathfinding problems with Answer Set Programming}
\author[Adrian Salewsky]{Adrian Salewsky}
\institute{University of Potsdam}
\date{00.00.2022}

\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{Table of Contents}
\tableofcontents
\end{frame}

\section{Introduction} 
\begin{frame}
\frametitle{Introduction}
\begin{itemize}
\item<2-> 
\end{itemize}
\end{frame}

\section{Abstraction Methods}
\begin{frame}
\frametitle{Shortest Path}
\begin{itemize}
\item<2-> 
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Node Combining}
\begin{itemize}
\item<2-> 
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Reachable Nodes}
\begin{itemize}
\item<2-> 
\end{itemize}
\end{frame}

\section{Auxiliary Programs}
\begin{frame}
\frametitle{Auxiliary Programs}
\end{frame}

\section{Benchmarking}
\begin{frame}
\frametitle{Benchmarking}
\end{frame}

\section{Conclusion}
\begin{frame}
\frametitle{Conclusion}
\begin{itemize}
\item<2-> 
\end{itemize}
\end{frame}

\end{document}